Data Analysis
================================

#Assignment 1


Download File
```{r, results = "hide"}
fileUrl <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile="./data.zip")
unzip("data.zip")
activity <- read.csv("activity.csv")
```

Transform Date
```{r, results = "hide"}
activity$date <- as.Date(activity$date)
```

Histogram of Steps Per Day
```{r Histogram of Steps Per Day}
hist(tapply(activity$steps, activity$date, sum), col = "green", 
     main = "Histogram of Steps Per Day", xlab = "Steps Per Day")
```

Mean and Median of Steps Per Day
```{r}
mean(tapply(activity$steps, activity$date, sum), na.rm=TRUE)
median(tapply(activity$steps, activity$date, sum), na.rm=TRUE)

```

Average Daily Activity Pattern, 
Time Series Plot of Daily Activity Pattern
```{r Daily Activity Pattern}
library(ggplot2)
ggplot(aes(x = interval, y = steps), data = activity) + geom_line(stat = 'summary', fun.y = mean, na.rm = TRUE)

```

Intervals sorted from highest average value to lowest average value, the 835th interval
has the highest value
```{r}
maxinterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
sort(maxinterval, decreasing = TRUE)
```

Calculate the # of missing values in the data set
```{r}
sum(is.na(activity$steps))
```

Fill in missing values using the mean of the interval
```{r, results = "hide"}
library(plyr)
library(Hmisc)
activity2 <- ddply(activity, "interval", mutate, imputed.value = impute(steps, mean))
activity2 <- activity2[order(activity2$date, activity2$interval),]
#imputed value converted to a numeric value
activity2$imputed.value <- as.numeric(activity2$imputed.value)
```

Histogram of Steps Per Day With Imputed Data
```{r Histograms of Setps Per Day with Imputed Data}
hist(tapply(activity2$imputed.value, activity2$date, sum), col = "green", 
     main = "Histogram of Steps Per Day", xlab = "Steps Per Day")
```

Mean and Median of Steps Per Day, the values are very similar to the original values
This does not change the prediction of the number of steps taken per day
```{r}
mean(tapply(activity2$imputed.value, activity$date, sum), na.rm=TRUE)
median(tapply(activity2$imputed.value, activity$date, sum), na.rm=TRUE)

```

Differences in activity patterns between weekdays and weekends, Create a variable for weekend and weekdays
```{r,results = "hide"}
##Convert the date to weekdays
activity2$wkday <- weekdays(activity2$date)
##Create weekend vector
weekend <- c("Saturday", "Sunday")
##create logical variable to determine if a day is a weekday or weekend
activity2$IsWeekend <- activity2$wkday %in% weekend
##change the logical variable to a factor variable
activity2$IsWeekend <- as.factor(activity2$IsWeekend)
##change the labels to weekday and weekend
levels(activity2$IsWeekend) <- c("Weekday", "Weekend")
```

```{r Weekend vs Weekday Time Series Plot}
ggplot(aes(x = interval, y = imputed.value), data = activity2) + 
  geom_line(stat = 'summary', fun.y = mean, na.rm = TRUE) +
  ylab("Number of Steps") + xlab ("Interval")+
  facet_wrap(~IsWeekend, nrow = 2)
```

